---
title: "Capacity building activities- 2024"
#date: "`r Sys.Date()`"
author: "IOTC Secretariat"
output:
  officedown::rdocx_document:
        reference_docx: "../templates/rdoc_template.docx"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      tab.cap.pre = "Tab. ", 
                      tab.cap.sep = ". ", 
                      tab.cap.style ="Table Caption", 
                      tab.cap.fp_text = officer::fp_text_lite(italic = TRUE, bold = TRUE), 
                      fig.cap = TRUE, 
                      fig.cap.pre = "Fig. ", 
                      fig.cap.sep = ". ", 
                      fig.cap.style = "Caption label",
                      dpi = 300
                      )
```

```{r Headers, echo = FALSE}
bl_header = function(txt, position) {
  block_list(
    fpar(values = txt, fp_p = fp_par(text.align = position, border.bottom = fp_border()), fp_t = fp_text_lite(color = "black", bold = FALSE))
  )
}

sec_pr = prop_section(
  header_default = bl_header("Data section capacity building activities - 2024", "right"),
  header_first = bl_header("Data section capacity building activities - 2024", "right"),
  header_even = bl_header("Data section capacity building activities - 2024", "right"),
)
```

```{r, echo = FALSE, eval = TRUE}
block_section(sec_pr)
```

```{r TextStyleExample, echo = FALSE}
fp = fp_par(
  text.align = "center", 
  padding.bottom = 20, padding.top = 120, 
  border.bottom = fp_border())

ft = fp_text(shading.color='#EFEFEF', bold = TRUE)
```


# Introduction

As part of the data section activities, providing technical assistance or any support that a CPC needs to boost up their ability to report fisheries data as required by IOTC Resolutions that entail data reporting. Besides direct requests from CPCs for support, in some instance, requests could be from Science working group, which are part of the work plan and priorities to improve the data available related to the specify working group or all species. This paper provide an overview of the data section activities for the year 2024, furthermore an updates on data-related requests from other working parties.

# Background

## Data section activities

Data section of the IOTC Secretariat undertook various activities aiming at improving the quality of data reported by CPCs, particularly for fisheries catching IOTC Species. The activities involve mainly direct meetings with CPCs or other organisations with interest to the IOTC data. There are circumstances where the Secretariat noticed continuous decline in the quality of data from a CPC, whereby the Secretariat will request to work closely with the CPC improve the data.
In 2024, the data section participated in 14 main activities, broadly categories as technical assistance, workshop, and external meeting, either as in country missions, webinar or with CPC representatives at the Secretariat office[@table]. The capacity building activities by country section provide an overview and outcomes of each mission conducted in 2024.

## Data request from other working party meetings

With the aim of improving the quality and availability of the data for stock assessment of the IOTC species, the working parties, made specific request to working party on data collection (WPDCS) to provide or improve specify data, which in many instance involve liaising directly to CPCs for the requests. There are circumstances where the data section has to launch specify project or create special format to request the data. However, not all CPCs willingly provide the information. Besides the working parties, Scientific committee (SC), make follow-up requests to WPDCS. In the last SC session and from subsequent working parties,  16 requests were made to on data related issues (@tab), which are listed in detailed in Appendix xxxx. 

## Progress made on the recommendations of WPDCS19

During the WPDCS19, four recommendations were made to SC (@tabl), categorized under different, in additional to recommendations of the program of work of WPDCS19. This document review the progress made on the recommendations, and were possible, make alternative recommendations for the participants to endorse. The progress on the recommendations are listed in Appendix ###.


# Capacity building Activities by country


## Tanzania

### Overview of the mission

Tanzania had had several assistance with their data collection and reporting systems for two consecutive years. The technical mission was a follow-up of the mission conducted in 2023. Tanzania is one of the CPCs with low catch of IOTC species from the coastal fisheries, however, their industrial fisheries is extending. Although the catches are low, the data reported are inconsistent in terms of species and fisheries, moreover, not all the datasets are reported. 

In 2019, Tanzania launched a pilot project called eCAS, which is a mobile App for data collectors (enumerators or samplers) were record the data from the landing sites and directly uploaded the information in the database system. During our last visit, the Secretariat learnt of the continuation of the eCAS, which are upgraded and now extended to more landing sites. The system is better adapted to small scale fisheries, however for fisheries catching tuna, limited information are recorded. Additionally, the system is only for coastal fisheries, and not for industrial fisheries, which in term should be using logbooks to collect data.

Often the are conflicted information between data sets reported. The shark data available from Tanzania consisted of data collected by the samplers, and also from Wildlife Conservation Society (WCS). WCS is also working closely with Tanzania government on National Plans of Action (NPOA) for the shark and rays for Tanzania.

During the mission in 2024, the Secretariat learnt that Tanzania anticipate to conduct a frame survey and conduct a comprehensive revision of the data collection system. One of the main activities of the mission is to train the officers reporting data to the Secretariat on how to improve the quality of the data reported, taking into consideration the errors made in previous submissions. The officers were trained on how to extract the required data, compiled and how to incorporate the data in the forms. In this case, as the Secretariat are developing new reporting tools, the exercises focused on the use of the new reporting tools

### Outcomes and recommendations

 - With the increasing information available from WCS on sharks and rays, one of the recommendations, was for the government to consolidate the data from WCS with samples data on sharks, raised to total catch before before reporting to the Secretariat.
 
 - Continuation of the integration of the tuna  and tuna-like data reporting requirements in the Fisheries Information Platforms (FiPs, *formal e-CAS*)
 
 - Develop a full assessment and re-estimation of coastal fisheries catch data for previous years
 
 - Follow-up with industrial fisheries companies to submit the data as required for industrial fisheries


## Bangladesh

### Overview of the mission

The technical mission to Bangladesh was held following the request from Bangladesh to provide technical support to implement the IOTC Conservation and Management Measures (CMM), related to statistical data requirements. As Bangladesh, is one of the newest IOTC members, the mission activities consisted of meetings with several officers in various locations involve with data collections. Furthermore, visiting landing sites to view and understand the Bangladeshi fisheries. During the mission the Secretariat officers had of overview of the data collection and reporting systems in Bangladesh, the Marine Fisheries Survey Management Unit (MFSMU). MFSMU conducts several fisheries assessment activities, besides collecting data. These include stock assessment for various species, vessel-based survey, land-based survey data management and reporting, analysis of fishery-dependent and independent data, and preparation of information for various meetings, include IOTC scientific meetings. 

Due to the nature of the fisheries, Bangladesh faces challenges for the data collection system, which impeded the reporting of data.

 - Insufficient manpower for data collection in future
 - Inadequate/inappropriate/partial data from some districts (upazilas)
 - Lack of budget and logistics in upazilas for data collection
 - Misinterpretation of data collection forms


### Outcomes and recommendations

 - The Secretariat had a better understanding of the institutions involved in fisheries monitoring and management, consisting of technical staff and several infrastructures
 - The main fisheries catching tuna and tuna-like species are gillnet, set bag nets and mid-water trawlers, with several species not reported to the Secretariat
  - The Sustainable Coastal and Marine Fisheries Project (SCMFP), helped to improve the collection system, resulting in large variability in catches observed in recent years, hence indicating underestimated in former data available at the Secretariat and FAO.
  - Assisting Bangladesh in adopting the new IOTC reporting forms
  - Bangladesh needs to estimate catches  based on the calendar year, as oppose the using fiscal year
  - Follow-up on the status of the industrial trawlers, if should be listed under RAV. 
  - Promote the use of Calipseo platform to DOF, as an appropriate tool to manage the data collection and reporting for the tuna and tuna-like species
  - Work closely with DOF to build on the information collected on the species composition of the catch for each gear obtained through the SCMFP project to (i) revise the current IOTC estimation of the catches and (ii) assess the feasibility of re-estimating the historical time series of catches of tuna and tuna-like species caught in Bangladesh over the whole time period in collaboration with DOF.

## Mozambique

### Overview of mission

The mission in Mozambique followed the request for technical assistance to enhance their statistical capabilities and optimize the utilization of the electronic platforms for data submission, which will help to overcome the challenges faced in fully meeting the data reporting requirements. The last data technical mission in Mozambique was in 2016, resulting in some improvement for the data reporting. In recent years, the quality of reporting from Mozambique deteriorated, which is a concern as Mozambique is chartering industrial vessels, without clear understanding of the reporting requirements.

The focus of the data technical mission in Mozambique was to review their fisheries collection and processing systems. Prior to the in person mission, a preliminary webinar was held with fisheries officers engaged with fisheries data processing in Mozambique. The meetings highlighted the challenges faced, which hinder the reporting of data.

Mozambique launched a taskforce, comprising statistical officers, scientists and managers from fisheries departments and institutions to address the IOTC reporting requirements, including those related to data submission. Mozambique fisheries vary by fishing techniques, and regionally , and not all fisheries necessary target tuna and tuna-like species. As a coastal state with large marine areas, fish consumption is very important for the local community, whereby tuna species only accounted for around %. The low contribution of the tuna fisheries in the Mozambique means low effort are put in data collection of tuna and tuna-like species. Similar to other countries, Mozambique is trying to move towards e-collection of data on landing sites. 
Currently, the collection of data are via enumerators and samples using paper format, which are then entered in the database systems. with the help of other institutions, Mozambique is trying to adopt a common data collection system use by several countries, [OPEN ARTFISH](https://www.fao.org/fishery/en/knowledgebase/152), to improve the digitalisation  process of artisanal fisheries, noting that Open ArTfish has limit in the type of data that are collected and processed, particular for the collection of biological sampling data. To conpensate for the lack of biological data, the fisheries department rely on Oceanographic Institute of Mozambique (InOM).

Mozambique, although is using the reporting form to submit the data, data reporting are often incomplete and missing key information. The Secretariat officers activities, also comprised training the officers on how to fill in the IOTC forms, particularly the main data reporting requirements, retained catches, catch and effort, and size frequency. As there are industrial fisheries in Mozambique, both purse seine and longline fisheries, the training also involve data reporting requirements of the industrial fisheries. It was challenging to fully complete the forms as the databases lack some of the main data requirements for the coastal fisheries, and for the industrial fisheries, the department is not receiving all the information from the fisheries companies.

To better understand the target species by fishery, the fisheries, and sampling systems, several landing sites were visited. Noting that targeted species are region-wise. South of Mozambique, consistin of very small fishing boats, target substantial demersal and crustacean species for local markets. Whereas in the North, the semi-industrial and small scale fisheries target small tuna and tuna-like species, with occasional catch of large pelagic species. It is noteworthy that sampling activities are limited on the sites, and no tools were observed from samplers.  

### Outcomes and recommendations

 - A comprehensive understanding of the institutions involved in the management and monitoring of fisheries in Mozambique, encompassing the main segments of the fisheries sector and their respective data collection systems.
 - Updates on the ongoing database management platform, with potential use of mobile App as mode of data collection
 - Follow-up with Mozambique on use on the various data reporting tools, and how they can improve reporting obligations
 - Encouraging Mozambique to participate more actively in the IOTC scientific meetings, upcoming regional training on species identification, to improve size frequency data report. 


## Indonesia

### Overview of mission

Indonesia catches is very important in the Indian Ocean, with significant catches of several IOTC species. The complexity of Indonesian fisheries, hinder the collection of fisheries data resulting into poor quality of reporting. In 2024, the Secretariat held several webinars with Indonesia and two in-person missions in Indonesia. The technical support to Indonesia focused on the re-estimation of the catch data of the entire fisheries, both coastal and industrial. For the reasons that (i) all fisheries operate within the NJA, (ii) retained catches are estimated broadly for all fisheries, (iii) changes in catches of one fishery categories, affect the catches of all fisheries, and (iv) any modification in estimation strategies in particular years, will lead to inconsistencies for other years. The missions in 2024 is triggered by the data presented during WPDCS19 by Indonesia, illustrated several inconsistencies, which could be attributed to gaps in the data collection and/or processing systems over time.

The mission in Indonesia involve reviewing the approach used by Indonesia to address the issues identified in the historical catch estimates. Moreover, help in developing a methodology to re-estimate catches prior to 2010. During the mission, the Secretariat had to review the format used for re-estimation, level of coverage of the landing catch from logbook and manual data collections, and furthermore, structure of the fleet use to categorize the fisheries. The exercise is hindered by limited past data collection, or scattering of repositories and processing of the data


###  Outcomes and recommendations

 -  Indonesia is to providing feedback on revision of historical catch estimates

## Sri Lanka

### Overview of mission

Since 2014, Sri Lanka made significant progress in data collection and reporting systems, improving the quality of the data reporting. However, for several, the Secretariat had provide excel templates for data reporting, which have several limitations, and from old excel version. In line to improve the reporting of data, the Secretariat introduced several reporting tools and new reporting templates recently, trained fisheries officers from CPCs on the utilization of the tools and completion of the templates. However, not all CPCs were able to capture to know-how of completing the new formats with associated tools, especially those with several fishery categories. 
During the IOTC Session in May 2024 (S28), Sri Lanka, clearly indicated that they could not approve the use the new templates in immediately as the staff will need time to better understand the processes. On this call, the Secretariat conduct a support mission in Sri Lanka, to help the fisheries officers to understand the new reporting processes and utilization of the new templates. Besides, focusing on improving data reporting, assessing the status of the Electronic Monitoring (EM) pilot project in Sri Lanka, and reviewing the gillnet fisheries data collection for conducting CPUE analysis, were part of the mission. 

The Secretariat had an overview of the data collection and processing systems of Sri Lanka, which are done by the Department of Fisheries and Aquatic Resources ([DFAR](https://www.fisheriesdept.gov.lk/) and of the National Aquatic Resource Research and Development Agency ([NARA](http://www.nara.ac.lk/)). Review of the historical data of tuna and tuna-like species of Sri Lanka was presented by the Secretariat, with the hope of improving the historical reported data.

The mission focused mainly on training the fisheries officers on the reporting tools, introduction of the online data reporting guidelines. A comparative analysis of the new and old reporting templates was presented showing the feasibility of the new templates and added tools to validate the data. 

The second component, CPUE data requirements, did not work as planned. The Secretariat hired a consultant, Dr. Paul Medley to provide assistance to Sri Lanka in improving the CPUE analysis presented, however, it was not possible to access the data to perform the exercise. 

The third component, EM project, the overview presented by Sri Lanka indicated failure of the project for several reasons: (i) challenges with installation due to vessels construction material; (ii) quick drainage of battery power, and (iii) difficulties in extract the collected data from the initial software use.

Visits to landing sites were crucial to view the various fisheries and type of fishery activities. The landing sites, and the activities of various vessel sizes, clearly denoted the various fisheries categories in Sri Lanka, and the importance of tuna and tuna-like species from each fishery. Samplers and enumerators are very active, whereby the collection of data either as port sampling or logbook are immediately transfer to central databases from the landing ports.

###  Outcomes and recommendations

The reporting tools and templates are new, and therefore faced some few issues, which the Secretariat will need to fix to improve the system. Furthermore, there are some of the fisheries or gear activities that are not encompassed by the new systems, which limited the reporting of some fisheries. The Secretariat is required to review the codes from various tools to ensure all fisheries and characteristics are captured, not only for Sri Lanka, but also from other CPCs.
There remains uncertainty about whether the consultant will obtain access to the data or if Sri Lanka will permit the use of this data for CPUE indices through alternative means. In relation to the use of EM in Sri Lankan fisheries, identifying cost-effective and reliable materials is crucial, and the IOTC should continue to support Sri Lanka in these efforts.

# Workshop


# Conclusion

During 2024, the data section undertook several activities, particular capacity building activities to improve the data reporting quality

